---
title: "db2pq: export PostgreSQL and WRDS data to Parquet"
format: gfm
---

I created `db2pq` as an R package for moving data from PostgreSQL into Apache Parquet files.
I believe that Parquet files are the best file format for researchers and data analysts who want "local" data, but without the hassle of managing a relational database server.
The design of the `db2pq` R package is modelled on that of [my Python package](https://pypi.org/project/db2pq/) of the same name.
However, I have elected to keep the R package focused on a PostgreSQL-to-Parquet pathway.
If you want to get Parquet data into a PostgreSQL database or move data from one PostgreSQL database to another, then look at [the Python package](https://pypi.org/project/db2pq/) for those pathways.

This package is designed both for general PostgreSQL sources and for the WRDS PostgreSQL service.

## What does `db2pq` do?

Some examples of what `db2pq` does are:

- Exports a single PostgreSQL table to Parquet.
- Exports all tables in a PostgreSQL schema to Parquet.
- Updates WRDS Parquet files only when the source table is newer.
- Reads and manages `last_modified` metadata embedded in Parquet files.
- Archives and restores different versions of Parquet files.

## Installation

To install from CRAN:

```r
install.packages("db2pq")
```

Or use `pak` to get the version from GitHub:

```r
# install.packages("pak")
pak::pak("iangow/db2pqr")
```

## Quickstart

### Load the package

Once you have installed `db2pq`, you can load into R it with `library()`.

```{r}
library(db2pq)
```

### Update a WRDS table

The main WRDS-related function is `wrds_update_pq()`.

```r
wrds_update_pq("dsi", "crsp", wrds_id = "your_wrds_id")
```


While you can pass a WRDS username directly to `wrds_update_pq()`, for general use, you should configure the WRDS username using *environment variables*.
The [authentication page](https://iangow.github.io/db2pqr/articles/authentication.html) explains how to do this.
The WRDS examples below assume that such a setup is in place.

Note that `wrds_update_pq()` only downloads data if WRDS has newer data than that stored in a local Parquet file.

```{r}
wrds_update_pq("dsi", "crsp")
```

### Force a re-download

If you want to overwrite existing data, you can use `force = TRUE`.

```{r}
wrds_update_pq("dsi", "crsp", force = TRUE)
```

### Using SAS metadata for updates

The update logic of `wrds_update_pq()` works by comparing information stored by WRDS in the PostgreSQL table comment with that retrieved and stored from previous runs of `wrds_update_pq()`. 
However, in some cases the PostgreSQL table comments are missing or may not provide the signal of "freshness" that we want to use.

WRDS data tables generally come in two forms: PostgreSQL tables and SAS files. 
The `wrds_update_pq()` function allows a user to specify `use_sas = TRUE`, which causes the function to check the "last updated" metadata from the corresponding WRDS SAS file instead.

In some cases, the schema used for SAS differs from that used for PostgreSQL and `wrds_update_pq()` can handle such cases using the `sas_schema` argument. 
For example, the WRDS PostgreSQL schema and the SAS schema do not line up for RavenPack and the following code handles this.

```{r}
wrds_update_pq("rpa_entity_mappings", "ravenpack_common",
               use_sas = TRUE, sas_schema = "rpa")
```

Additional setup is needed for this SAS metadata path; see the [WRDS SSH setup article](https://iangow.github.io/db2pqr/articles/wrds-ssh.html) for details.

### Update all tables in a schema

```{r}
wrds_schema_to_pq("ff", views = TRUE)
```

### Export a custom WRDS SQL query

```r
wrds_sql_to_pq(
  "SELECT permno, date, ret FROM crsp.dsf WHERE date >= '2024-01-01'",
  table_name = "dsf_recent",
  schema = "crsp"
)
```

### Export a local PostgreSQL table

```r
db_to_pq(
  table_name = "company",
  schema = "comp",
  keep = c("gvkey", "conm"),
  rename = c(conm = "company_name")
)
```

### Check when local Parquet files were last updated

```{r}
pq_last_modified(schema = "crsp")
```

## Parquet layout

Files are organized as:

```
<DATA_DIR>/<schema>/<table>.parquet
```

For example:

```
~/pq_data/crsp/dsi.parquet
```

The `DATA_DIR` environment variable sets the root directory. It can also be
passed directly as `data_dir` to any function.

When `archive = TRUE`, replaced files are moved to:

```
<DATA_DIR>/<schema>/archive/<table>_<timestamp>.parquet
```

## License

MIT License. See `LICENSE.md`.
