---
title: "WRDS authentication"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

There are two elements required for authentication of a WRDS PostgreSQL connection:

1. the WRDS username
2. the WRDS password for that username

One goal of the `db2pq` package is to keep scripts used to manage WRDS data free of hard-coded credentials.
In addition, the `db2pq` package endeavours to build on existing credential management processes that users may have.

## Resolving the WRDS username

For credential helpers, `db2pq` resolves a WRDS username using the following order:

1. an explicit `wrds_id` argument supplied to the function
2. the `WRDS_ID` environment variable
3. the `WRDS_USER` environment variable^[`WRDS_USER` is accepted mainly for compatibility with the Tidy Finance setup.]
4. the keyring username created by `wrds::wrds_set_credentials()`

Because of this order, an explicit function argument can be used to override environment variables or keyring defaults:

```{r, eval = FALSE}
wrds_update_pq("company", "comp", wrds_id = "your_wrds_id")
```

The `wrds_get_username()` function can be used to see the resolved WRDS username:

```{r, eval = FALSE}
wrds_get_username()
```

If `wrds_get_username()` does not resolve to a WRDS username using the process described above, then the user is prompted to supply one.
The supplied username is used for the current session and the user is given the option to save the username for future R sessions.

## Resolving the WRDS password

Once `db2pq` has resolved a WRDS username, it needs a password for that username to connect to the WRDS PostgreSQL database.

Passwords are resolved in the following order:

1. an explicit `password` argument, where the function supports one
2. a matching entry in PostgreSQL's `.pgpass` password file
3. an existing `WRDS_PASSWORD` environment variable^[`WRDS_PASSWORD` is
   read for compatibility with the Tidy Finance setup.
   `db2pq` does not create or persist this environment variable.]
4. an interactive password prompt when no matching `.pgpass` entry or
   `WRDS_PASSWORD` is found^[When prompting is disabled and `db2pq` passes no
   password to the connection, PostgreSQL/libpq may still use another
   credential mechanism, such as `PGPASSWORD`.]

If the username can only be resolved from credentials created by
`wrds::wrds_set_credentials()`, `db2pq` falls back to `wrds::wrds_connect()`.
That route lets the `wrds` package use its keyring credentials.

The `wrds_check_credentials()` function can be used to test the resolved WRDS username and password:

```{r, eval = FALSE}
wrds_check_credentials()
```

If an interactive `db2pq` connection prompts for a password and the connection succeeds, the user-supplied password is saved to `.pgpass` for future connections.
When `db2pq` uses an existing `WRDS_PASSWORD` because `.pgpass` has no matching entry, it saves that password to `.pgpass` after a successful connection too.

An explicit password can also be checked without saving it:

```{r, eval = FALSE}
wrds_check_credentials(password = "your_wrds_password", save = FALSE)
```

PostgreSQL's `.pgpass` mechanism is preferred to putting a password in R code.
It also works with RPostgres/libpq and other PostgreSQL tools, not only
`db2pq`.

The default PostgreSQL password-file path is:

- `~/.pgpass` on macOS and Linux
- `%APPDATA%/postgresql/pgpass.conf` on Windows

A WRDS `.pgpass` entry has the PostgreSQL password-file shape:

```text
wrds-pgdata.wharton.upenn.edu:9737:wrds:your_wrds_id:your_wrds_password
```

On macOS and Linux, PostgreSQL expects the file to be readable only by the owner.
Password entries saved by `db2pq` use that permission automatically;
for a file created manually, set it with:

```sh
chmod 600 ~/.pgpass
```

## Inspecting and testing your setup

Use `wrds_check_credentials()` for a live connection test:

```{r, eval = FALSE}
wrds_check_credentials()
```

If the matching WRDS entry is not already in `.pgpass`, an interactive
`wrds_check_credentials()` call asks for the WRDS PostgreSQL password. When the connection test succeeds, it writes that password to `.pgpass` for later libpq-backed connections.
You can pass `password = ...` explicitly for a non-interactive setup check, use `save = FALSE` for a one-time check, or use `passfile = ...` to target a specific PostgreSQL password file.

The returned diagnostic reports whether a small WRDS PostgreSQL connection succeeded and whether the matching target appears in `.pgpass`.

## Existing `wrds` package credentials

If you already use the `wrds` package, its keyring-based setup remains a
valid path:

```{r, eval = FALSE}
wrds::wrds_set_credentials()
wrds_update_pq("company", "comp")
```

Use this route when you want to stay close to existing R WRDS code you may have using the `wrds` package.
Use the environment-variable plus `.pgpass` route when you want credentials to work across `db2pq`, RPostgres/libpq, and other PostgreSQL tools on the same machine.

## SAS-based workflows

Most `db2pq` WRDS work uses WRDS PostgreSQL.
If you set `use_sas = TRUE`, `db2pq` also retrieves metadata from the WRDS SAS side over SSH.
That SSH setup is separate from PostgreSQL password handling; see the
[WRDS SSH setup article](wrds-ssh.html) before using SAS metadata in
unattended refresh scripts.

## Environment variables in R

`Sys.setenv()` is useful for a running R session:

```{r, eval = FALSE}
Sys.setenv(
  WRDS_ID = "your_wrds_id",
  DATA_DIR = "~/Dropbox/pq_data"
)
```

For settings you want available when R starts, use `.Renviron`.
This is the closest R analogue to a simple environment-variable file for this workflow.
R searches for `.Renviron` in the current working directory and then in the user's home directory, so there are two useful patterns:

- `~/.Renviron`: user-level defaults that apply broadly on your machine
- `.Renviron` in a project directory: project-local values for that project

A user-level `~/.Renviron` is a good place for a default WRDS username or
default Parquet repository:

```text
WRDS_ID=your_wrds_id
DATA_DIR=~/Dropbox/pq_data
```

A project-local `.Renviron` is useful when one project should override the user-level defaults without changing other R work on the same machine.
For example, it can select a project-specific Parquet repository:

```text
DATA_DIR=~/research/project-data/pq_data
```

The environment file should not be committed if it contains private,
machine-specific, or secret values.
After editing a `.Renviron` file, start a new R session or load it explicitly with `readRenviron()`.

For first-time `DATA_DIR` setup, call a Parquet helper such as `wrds_update_pq()` in an interactive session and let its default data-directory handling run.
In an interactive session without a `DATA_DIR`, `db2pq` asks for a data directory, can create a new directory if needed, and offers to store `DATA_DIR` in either project-level `.Renviron` or user-level `~/.Renviron`.
That scope choice is useful for data repositories; the interactive WRDS username prompt stores `WRDS_ID` in user-level `~/.Renviron`.

## Related pages

- WRDS to Parquet article
- Data management article
- `wrds_get_username()`
- `wrds_check_credentials()`
