---
title: "Data Management for Research"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Management for Research}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)
pq_path <- function(schema, table) {
  file.path(Sys.getenv("DATA_DIR", unset = ""), schema, paste0(table, ".parquet"))
}
has_pq <- function(schema, table) {
  nzchar(Sys.getenv("DATA_DIR", unset = "")) && file.exists(pq_path(schema, table))
}
live_repo <- all(
  identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_pkg("arrow"),
  has_pkg("dplyr"),
  has_pq("crsp", "dsi"),
  has_pq("comp", "company")
)
doc_data_dir <- file.path(tempdir(), "db2pq-data-management")
if (live_repo) {
  unlink(doc_data_dir, recursive = TRUE)
  dir.create(file.path(doc_data_dir, "comp"), recursive = TRUE, showWarnings = FALSE)
  file.copy(
    pq_path("comp", "company"),
    file.path(doc_data_dir, "comp", "company.parquet"),
    overwrite = TRUE
  )
}
```

This article adapts a broader set of data-management ideas for the narrower
task that `db2pq` is designed to support: building and maintaining a local
Parquet repository from WRDS or another PostgreSQL database.

The goal is to separate three concerns that often get mixed together in
research code:

1. Getting data from the source system.
2. Storing a reusable local copy in an efficient format.
3. Running analysis code against that local copy.

`db2pq` focuses on the first two steps. Analysis code can then use packages
such as `arrow`, `duckdb`, `dplyr`, or `polars` to work with the Parquet files.

## Principles

Good research data management is mostly about making later work boring:
scripts should be rerunnable, file locations should be predictable, and it
should be clear which files came from an external source and which files were
created by project code.

Several principles are useful:

1. Keep source data immutable where practical.
2. Make derived data reproducible from code.
3. Use stable, documented file locations.
4. Record enough metadata to know when a file was created and what source it
   represents.
5. Avoid putting large, private, or licensed data files in Git.

`db2pq` is not a full project-management framework. It supplies a narrow piece
of infrastructure: a repeatable way to create local Parquet copies of database
tables and to inspect the metadata attached to those copies.

## Prefer Portable File Formats

Research data often outlives the first tool used to analyze it. Parquet is a
useful storage format because it is columnar, compressed, typed, and readable
from many systems. A local Parquet repository can be queried from R, Python,
DuckDB, Arrow, Spark, and other tools without repeatedly connecting to WRDS.

This is especially useful when a project uses more than one language. A refresh
script can use `db2pq` to create files, while later analysis can read the same
files from R or Python.

## Use a Stable Data Directory

Set `DATA_DIR` once per project or machine so that scripts can refer to a
logical data repository without hard-coding paths throughout the analysis.

```{r, eval = FALSE}
Sys.setenv(DATA_DIR = "~/Dropbox/pq_data")
```

Files created by `db2pq` are stored using the layout:

```text
<DATA_DIR>/<schema>/<table>.parquet
```

For example:

```text
~/Dropbox/pq_data/crsp/dsi.parquet
~/Dropbox/pq_data/comp/company.parquet
```

## Treat Downloads as a Refresh Step

The scripts that download data should usually be separate from the scripts that
analyze data. A refresh script might contain calls like these:

```{r, eval = FALSE}
library(db2pq)

wrds_update_pq("dsi", "crsp")
wrds_update_pq("company", "comp")
wrds_update_pq("funda", "comp", keep = c("gvkey", "datadate", "at", "sale"))
```

Because `wrds_update_pq()` compares WRDS metadata with metadata embedded in the
local Parquet file, it can be rerun without downloading a table that is already
current.

The same idea applies to non-WRDS PostgreSQL sources. Put source-system refresh
logic in one place, then let analysis code consume the resulting files.

## Keep Raw and Derived Data Separate

A useful convention is to keep source-table exports in one repository and write
derived research datasets somewhere else. For example:

```text
~/Dropbox/pq_data/          # Local source-table repository
data/derived/               # Project-specific derived files
```

The source-table repository should be reproducible from WRDS or PostgreSQL.
Derived files should be reproducible from project code.

This separation matters because source data and derived data have different
rules. Source exports should be refreshed deliberately and should preserve the
database table as closely as is useful. Derived files can be redesigned as the
research design changes.

## Make Transformations Explicit

Avoid hiding important filters, joins, and sample restrictions inside ad hoc
interactive work. Put them in scripts or functions. A useful pattern is:

```text
source database -> local Parquet repository -> derived research data -> tables/figures
```

`db2pq` can help with the first arrow. The later arrows should usually live in
project-specific code, where sample construction choices are visible to readers
and coauthors.

## Record Metadata

Parquet files written by WRDS helpers include `last_modified` metadata when the
source table provides it. Use `pq_last_modified()` to audit a repository:

```{r, eval = FALSE}
pq_last_modified(schema = "crsp")
pq_last_modified(schema = "comp")
```

For data-management work, this is often more useful than relying only on local
file modification times, because the metadata records what `db2pq` knows about
the source table.

Metadata will not answer every provenance question, but it gives a lightweight
audit trail. For more complex projects, combine it with project logs, scripts,
or a data dictionary that records table sources and refresh dates.

When this article is rendered with access to a local Parquet repository, the
following chunk shows the embedded source metadata for a real WRDS-derived file.

```{r live-metadata-dsi, eval = live_repo}
Sys.getenv("DATA_DIR")
pq_last_modified(table_name = "dsi", schema = "crsp")
```

## Archive Before Replacing

For tables where changes matter, use `archive = TRUE` so an existing file is
moved aside before it is replaced:

```{r, eval = FALSE}
wrds_update_pq("company", "comp", archive = TRUE)
pq_last_modified("company", "comp", archive = TRUE)
```

Archives are stored under:

```text
<DATA_DIR>/<schema>/archive/
```

This gives you a simple local history without turning the data repository into a
Git repository.

Archiving is most useful for relatively small reference tables or for cases
where source revisions may affect published results. For very large tables,
keeping every historical copy may be too expensive; in those cases, record the
refresh date and source metadata instead.

To make the archive mechanics concrete without modifying the main local data
repository, this article copies `comp.company` into a temporary documentation
repository and archives/restores that copy.

```{r live-archive-copy, eval = live_repo}
pq_last_modified(table_name = "company", schema = "comp", data_dir = doc_data_dir)
```

Archiving moves the active file into the schema's `archive` directory.

```{r live-archive-run, eval = live_repo, message = FALSE}
archived_company <- pq_archive("company", "comp", data_dir = doc_data_dir)
basename(archived_company)

pq_last_modified(table_name = "company", schema = "comp",
                 data_dir = doc_data_dir, archive = TRUE)[
  c("file_name", "last_mod")
]
```

Restoring the archived file makes it the active copy again.

```{r live-restore-run, eval = live_repo, message = FALSE}
pq_restore(tools::file_path_sans_ext(basename(archived_company)), "comp",
           data_dir = doc_data_dir, archive = FALSE)

pq_last_modified(table_name = "company", schema = "comp", data_dir = doc_data_dir)
```

## Keep Large Data Out of Git

Code, documentation, and small test fixtures belong in Git. Full WRDS exports
generally do not. If a package or project needs example data, use a small,
non-sensitive file and place it deliberately in `inst/extdata` or another
documented location.

For package documentation, build examples so they can run without private WRDS
credentials. Longer examples can be shown with `eval = FALSE`, while the local
rendered website can still refer to small local Parquet fixtures when those are
safe to publish.

## Plan for Collaboration

Shared data workflows should minimize assumptions about one person's machine.
Use environment variables, relative project paths, and documented setup steps
instead of absolute paths such as `/Users/name/...`.

For WRDS credentials, prefer setup that each user can reproduce locally:

```{r, eval = FALSE}
Sys.setenv(WRDS_ID = "your_wrds_id")
wrds_check_credentials()
```

Do not store credentials in project scripts. The package helpers are designed to
work with environment variables, `.pgpass`, and the credential mechanisms used
by the `wrds` package.

## Suggested Project Pattern

A compact project structure might look like this:

```text
analysis-project/
  R/
  scripts/
    refresh-data.R
  data/
    derived/
  reports/
  renv.lock
```

The refresh script owns calls to `db2pq`. Analysis scripts read Parquet files
from `DATA_DIR` and write project-specific outputs to `data/derived`.

## Live Examples with Local Parquet Files

The examples in this section are designed for the locally rendered
documentation site. They run when `DATA_DIR` points to a Parquet repository that
contains the relevant WRDS-derived files. If those files are not present, the
chunks are skipped.

```{r local-example-setup, include = FALSE}
live_crsp <- all(
  identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_pkg("DBI"),
  has_pkg("duckdb"),
  has_pkg("dplyr"),
  has_pkg("dbplyr"),
  has_pkg("ggplot2"),
  has_pq("crsp", "dsf"),
  has_pq("crsp", "stocknames")
)
```

For example, if your repository contains `crsp.dsf` and `crsp.stocknames`, you
can query the files directly with DuckDB without importing the full data into R.

```{r live-crsp-connect, eval = live_crsp}
library(DBI)
library(dplyr)
library(dbplyr)
library(ggplot2)

db <- dbConnect(duckdb::duckdb())

parquet_tbl <- function(con, schema, table) {
  path <- normalizePath(pq_path(schema, table), winslash = "/", mustWork = TRUE)
  view <- paste(schema, table, sep = "_")
  DBI::dbExecute(
    con,
    paste0("CREATE OR REPLACE VIEW ", view, " AS SELECT * FROM read_parquet('", path, "')")
  )
  tbl(con, view)
}

dsf <- parquet_tbl(db, "crsp", "dsf")
stocknames <- parquet_tbl(db, "crsp", "stocknames")
```

A first check is simply to count the rows in the daily stock file.

```{r live-crsp-count, eval = live_crsp}
dsf |>
  count() |>
  collect()
```

Because DuckDB can query Parquet files in place, summaries can be computed
without first reading every column into memory.

```{r live-crsp-year-count, eval = live_crsp}
dsf |>
  mutate(year = as.integer(strftime(date, "%Y"))) |>
  count(year) |>
  arrange(desc(year)) |>
  collect()
```

Filtering to one trading day reads only the rows and columns needed for that
query.

```{r live-crsp-one-day, eval = live_crsp}
dsf_subset <-
  dsf |>
  filter(date == as.Date("1986-01-07")) |>
  select(permno, date, ret, prc, vol) |>
  collect()

head(dsf_subset)
```

The same files can be used for a small analysis. Here we identify Apple's CRSP
`permno` from `stocknames`, then compute a cumulative return series from
`dsf`.

```{r live-crsp-apple, eval = live_crsp, fig.width = 7, fig.height = 4}
apple_permno <-
  stocknames |>
  filter(grepl("^APPLE COM", comnam)) |>
  distinct(permno) |>
  collect() |>
  pull(permno) |>
  min()

apple_returns <-
  dsf |>
  filter(permno == apple_permno) |>
  select(permno, date, ret) |>
  arrange(date) |>
  collect() |>
  mutate(cumret = exp(cumsum(log(1 + coalesce(ret, 0)))))

ggplot(apple_returns, aes(x = date, y = cumret)) +
  geom_line() +
  labs(x = NULL, y = "Cumulative return")
```

```{r live-crsp-disconnect, eval = live_crsp, include = FALSE}
dbDisconnect(db, shutdown = TRUE)
```

These examples use a local Parquet repository as the analysis substrate. The
refresh step that creates the files is still separate:

```{r, eval = FALSE}
wrds_update_pq("dsf", "crsp")
wrds_update_pq("stocknames", "crsp")
```

That separation is the main point. `db2pq` handles the database-to-Parquet
refresh; analysis code can then operate against local, typed, columnar files.
