---
title: "Choosing a DATA_DIR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`db2pq` writes Parquet files into a local repository. `DATA_DIR` names the
root of that repository so the same refresh code can work with a repository on
different machines or in different projects.

For example, this call:

```{r, eval = FALSE}
library(db2pq)

wrds_update_pq("dsi", "crsp")
```

writes the active Parquet file to:

```text
<DATA_DIR>/crsp/dsi.parquet
```

The standard layout is:

```text
<DATA_DIR>/<schema>/<table>.parquet
```

When updates use `archive = TRUE`, replaced files are moved under:

```text
<DATA_DIR>/<schema>/archive/<table>_<timestamp>.parquet
```

## First-Time Setup

All Parquet helpers use `db2pq_data_dir()` for their default `data_dir`.
When `DATA_DIR` is missing in an interactive session, that helper says so
before asking whether to open a directory chooser. You can select a directory
there, or decline the chooser and enter a path in R. If an entered path does
not exist, the helper can create it after confirmation.

```{r, eval = FALSE}
db2pq_data_dir()
```

The selected directory is available as `DATA_DIR` for the current R session.
The helper then offers three persistence choices:

- project `.Renviron`
- user `~/.Renviron`
- no persistent setting

After editing or creating `.Renviron`, start a new R session before expecting
R to read the new value automatically.

## Choosing the Scope

A user-level `DATA_DIR` is appropriate for a long-lived local Parquet library
that many projects use. For example, `~/.Renviron` might contain:

```text
DATA_DIR=~/Dropbox/pq_data
```

A project-level `DATA_DIR` is useful when one project should use a frozen or
separate repository. A `.Renviron` file in that project might contain:

```text
DATA_DIR=data/pq_data
```

R reads project `.Renviron` values when the project session starts, so this
lets project code keep using the same calls while pointing at a different
Parquet repository.

Do not commit `.Renviron` or Parquet files when they contain private,
machine-specific, or licensed data.

## Explicit Overrides

You do not have to use `DATA_DIR`. An explicit `data_dir` argument overrides
the default for a call:

```{r, eval = FALSE}
wrds_update_pq("dsi", "crsp", data_dir = "data")
```

That pattern is useful when a refresh script should write a project-level copy
while your default `DATA_DIR` still points at a shared local library.

For non-interactive code, set `DATA_DIR` or pass `data_dir` explicitly when the
working directory is not the repository root you intend to use.

## Related Pages

- [Authentication](authentication.html) covers `WRDS_ID`, passwords, and
  `.pgpass`.
- [WRDS to Parquet](wrds-to-parquet.html) shows the refresh workflow.
- [Data Management for Research](data-management.html) discusses shared
  repositories, project-level copies, and archives in a broader research
  workflow.
