---
title: "Creating a Parquet data repository"
bibliography: references.bib
---

```{r}
library(db2pq)
```

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)
pq_path <- function(schema, table) {
  file.path(Sys.getenv("DATA_DIR", unset = ""), schema, paste0(table, ".parquet"))
}
has_pq <- function(schema, table) {
  nzchar(Sys.getenv("DATA_DIR", unset = "")) && file.exists(pq_path(schema, table))
}
live_repo <- all(
  identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_pkg("arrow"),
  has_pkg("dplyr"),
  has_pq("crsp", "dsi"),
  has_pq("comp", "company")
)
doc_data_dir <- file.path(tempdir(), "db2pq-data-management")
if (live_repo) {
  unlink(doc_data_dir, recursive = TRUE)
  dir.create(file.path(doc_data_dir, "comp"), recursive = TRUE, showWarnings = FALSE)
  file.copy(
    pq_path("comp", "company"),
    file.path(doc_data_dir, "comp", "company.parquet"),
    overwrite = TRUE
  )
}
```

This article outlines some ideas related to the task that `db2pq` is designed to support: 
building and maintaining a local Parquet repository of data from WRDS and other sources.

The goal is to separate three concerns that often get mixed together in
research code:

1. Getting data from the source system.
2. Storing data locally in an efficient format.
3. Running analysis code against local data.

The `db2pq` focuses on the first two steps. 
Analysis code can then use packages such as `arrow`, `duckdb`, `dplyr`, or `polars` to work with the Parquet files.

##  Some concepts in data management

### Scope

Many datasets are **project-specific** datasets, meaning that only have use within a single project (e.g., paper).
Examples of project-specific would include experimental data generated in a particular study.

Other datasets are **general-purpose** datasets, meaning that they contain data that might be relevant to many studies.
Classic examples in a business-school context would be the [US stock price files](https://iangow.github.io/far_book/identifiers.html#the-crsp-database) offered by the Center for Research in Security Prices, LLC (CRSP) or [financial statement data](https://iangow.github.io/far_book/fin-state.html#financial-statement-data) provided by Compustat, or economic time-series data provided by various statistical offices around the world.

Other datasets are **project-level** datasets, meaning that the particular data sets are somehow frozen for a particular project, even though the nature of the data otherwise puts them in the category of general-purpose datasets.
For example, I might want to fix on a particular version of `comp.g_funda`, Compustat's global dataset for annual financial statements for my project, even though this dataset has relevance beyond a specific project.^[As academic researchers generally get Compustat data through Wharton Research Data Services (WRDS), I refer to this dataset using the nomenclature used by WRDS. Here "global" means "not the United States (or Canada)".]

One reason for having project-level arises in the context of reproducibility.
If I have published a paper, then the replication package for that paper should ideally contain the data used to produce the exact results in the paper.
For this purpose, if the paper used `comp.g_funda` data, then the ideal replication package would include the precise *project-level* version of that data set used to produce the paper.
Of course, in reality, one cannot simply post the project-level version of `comp.g_funda` as part of a public replication package.
Nonetheless, the authors themselves should have a project-level version of the dataset that *they* retain.
This much aligns with the views of @Welch:2019aa, who suggests that "the author should keep a private copy of the full data set with which the results were obtained."

As far as WRDS data are concerned, the `db2pq` package aims to facilitate managing WRDS data either as general-purpose data or as project-level data.^[WRDS data are not project-specific data sets pretty much by definition.]
The  `db2pq` package can leverage the environment variable `DATA_DIR` to identify the root directory for the data repository in question.
By default, new WRDS data will go in the matching schema (i.e., subdirectory) of the directory indicated by `DATA_DIR`.

So, depending on how you want to manage data for a given project, you may set `DATA_DIR` to a **project**-level directory or to a **user**-level directory.



### Storage formats

While there are many storage formats available for data, I think a strong case can be made for Parquet being the default storage format for many users.
If you use R or Python, I think the case is easy to make.
Many software packages can read Parquet data and some of them (e.g., the `arrow` package or DuckDB) will absolutely fly with Parquet data.

I believe that recent editions of Stata can read Parquet files, though the way Stata operates means that Stata users are unlikely to see the performance benefits Parquet offers.^[Of course, if a user cared about performance with data manipulation, he probably wouldn't be using Stata to begin with.]
SAS users might find the case for Parquet less compelling, though there are probably benefits in moving to a storage medium that is more compact, less proprietary, and more likely to be supported in a few years' time.

Of course, an alternative to using Parquet files would be using a database, such as PostgreSQL.
I think such systems have a lot of merit (and I have used PostgreSQL to store WRDS data since 2011), but I think they are more complicated for most users' needs and their benefits (e.g., shared access to data and rock-solid assurance) are less meaningful for most.

Another alternative is the CSV file, perhaps compressed.
I think if one were sending data on the next Voyager mission, then CSV might be the chosen format.^[Each of the two Voyager spacecraft, launched by NASA in 1977, carry the Voyager Golden Record, a gold-plated copper phonograph record intended as a message to any intelligent extraterrestrial lifeforms that might encounter the probes.
If we wanted to give data to such lifeforms, I think it would be (quoted) CSV data and written on paper.]
Or if you really, really wanted data novices to inspect your data in Excel or Word, then CSV might be the go-to option.
Or perhaps you want to put your data in a written form in a book for users to type in.
For any other purpose with serious data needs, I think Parquet dominates.

One issue with CSV is that one is always dealing with type inference (string, integer, timestamp) and I think that type inference is one of those problems you want to solve once for any given dataset.
For the WRDS data that is the focus of this note, I think CSV is to be avoided.

## Setting `DATA_DIR`

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


### First-Time Setup

The root of the data repository is set in  `DATA_DIR`, so the same refresh code can work with a repository on
different machines or in different projects.
When `DATA_DIR` is missing in an interactive session, you will be asked to select a directory.
You will then be given three choices for persisting the directory you have chosen:

- a **project**-level `DATA_DIR` (stored in `.Renviron`)
- a **user**-level `DATA_DIR` (using  `~/.Renviron`)
- no persistence

`db2pq` writes Parquet files into a local repository.

For example, this call:

```{r, eval = FALSE}
library(db2pq)

wrds_update_pq("dsi", "crsp")
```

writes the active Parquet file to:

```text
<DATA_DIR>/crsp/dsi.parquet
```

The standard layout is:

```text
<DATA_DIR>/<schema>/<table>.parquet
```

When updates use `archive = TRUE`, replaced files are moved under:

```text
<DATA_DIR>/<schema>/archive/<table>_<timestamp>.parquet
```

### Explicit Overrides

You do not have to use `DATA_DIR`. 
An explicit `data_dir` argument overrides the default for a call:

```{r, eval = FALSE}
wrds_update_pq("dsi", "crsp", data_dir = "data")
```

That pattern is useful when a refresh script should write a project-level copy
while your default `DATA_DIR` still points at a shared local library.

For non-interactive code, set `DATA_DIR` or pass `data_dir` explicitly when the
working directory is not the repository root you intend to use.

## Related Pages

- [Authentication](authentication.html) covers `WRDS_ID`, passwords, and
  `.pgpass`.
- [WRDS to Parquet](wrds-to-parquet.html) shows the refresh workflow.
