---
title: "Parquet utility functions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_pq <- function(schema, table) {
  data_dir <- Sys.getenv("DATA_DIR", unset = "")
  nzchar(data_dir) &&
    file.exists(file.path(data_dir, schema, paste0(table, ".parquet")))
}
live_repo <- all(
  identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_pq("comp", "company"),
  has_pq("crsp", "dsi")
)
company_file <- file.path(
  Sys.getenv("DATA_DIR", unset = "."),
  "comp",
  "company.parquet"
)
```

This page groups together examples for the Parquet utility helpers:
`pq_data_dir()`, `pq_last_modified()`, `pq_archive()`, `pq_restore()`, and
`pq_remove()`.

This page shows how the utilities fit together in the practical workflow of maintaining a local Parquet repository.

## When to use these helpers

These functions generally help when you already have a Parquet repository and want to:

- inspect what is in it
- check which vintage of a table you currently have
- archive a current active file before replacing it
- restore an archived vintage
- remove files

The related API pages are:

- `pq_data_dir()`
- `pq_last_modified()`
- `pq_archive()`
- `pq_restore()`
- `pq_remove()`

## Setup

Load the packages used in the examples:

```{r parquet-utils-packages, message = FALSE}
library(db2pq)
library(dplyr)
```

Set `DATA_DIR` to the Parquet repository you want to inspect:

```{r, eval = FALSE}
Sys.setenv(DATA_DIR = "~/Dropbox/pq_data")
```

Use `pq_data_dir()` when you want to confirm which repository `db2pq` will use
by default:

```{r parquet-utils-data-dir, eval = FALSE}
pq_data_dir(prompt = FALSE)
```


The examples below are rendered from a local Parquet repository. The examples
that move files run only during a local `pkgdown` render when the repository
contains `comp.company` and `crsp.dsi`.

## Inspect a Schema Directory

Use `pq_last_modified()` without `table_name` to summarize the active Parquet
files in a schema:

```{r parquet-utils-list-schema, eval = live_repo}
pq_last_modified(schema = "comp") |>
  select(file_name, table, last_mod) |>
  head(10)
```

If you want to inspect archived files instead:

```{r parquet-utils-list-archive, eval = live_repo}
pq_last_modified(schema = "comp", archive = TRUE) |>
  select(file_name, table, last_mod) |>
  head(10)
```

If your project uses a repository outside the default `DATA_DIR`, pass
`data_dir` explicitly:

```{r parquet-utils-list-explicit-dir, eval = live_repo}
pq_last_modified(schema = "crsp", data_dir = Sys.getenv("DATA_DIR")) |>
  select(file_name, table, last_mod) |>
  head(5)
```

## Check the Current Active Vintage

With `table_name` and `schema`, `pq_last_modified()` returns the raw embedded
`last_modified` metadata string for the active file:

```{r parquet-utils-active-vintage, eval = live_repo}
pq_last_modified(table_name = "dsi", schema = "crsp")
```

This is often the fastest way to confirm what vintage a local Parquet file
represents before starting analysis.

## Inspect Archived Vintages

If you archive replaced files, you can ask for the archived versions of a
table:

```{r parquet-utils-archived-vintages, eval = live_repo}
pq_last_modified(table_name = "company", schema = "comp", archive = TRUE) |>
  select(file_name, table, last_mod, last_mod_str) |>
  tail(10)
```

That returns a table-like summary of the archived vintages for the requested
dataset. To inspect archived files for a whole schema, use `schema` without
`table_name`:

```{r parquet-utils-schema-archive-summary, eval = live_repo}
pq_last_modified(schema = "comp", archive = TRUE) |>
  select(file_name, table, last_mod) |>
  head(10)
```

## Archive the Currently Active File

You can archive a file manually even outside an update workflow:

```{r, eval = FALSE}
pq_archive(table_name = "company", schema = "comp")
```

Or archive an exact file path. During the live render, this moves the current
`comp.company` file into its archive directory:

```{r parquet-utils-archive-file, eval = live_repo, message = FALSE}
company_archive <- pq_archive(file_name = company_file)
basename(company_archive)
```

This is useful when you want to preserve the current active vintage before
running an experimental refresh or downstream transformation.

## Restore an Archived Vintage

To promote an archived file back into the active schema directory:

```{r parquet-utils-restore-file, eval = live_repo, message = FALSE}
restored_company <- pq_restore(
  tools::file_path_sans_ext(basename(company_archive)),
  "comp",
  archive = FALSE
)
basename(restored_company)
```

The archived basename may include or omit the `.parquet` suffix. If an active
destination file already exists, `pq_restore()` can archive that file first
with its default `archive = TRUE`.

```{r parquet-utils-restored-vintage, eval = live_repo}
pq_last_modified(table_name = "company", schema = "comp")
```

## Remove a File Explicitly

Use `pq_remove()` when you want to delete an active or archived file rather
than archive it. The removal examples below are shown but not run during the
documentation build. For an active file:

```{r, eval = FALSE}
pq_remove(table_name = "dsi", schema = "crsp")
```

To remove an archived file:

```{r, eval = FALSE}
pq_remove(
  table_name = "company_20260407T060000Z",
  schema = "comp",
  archive = TRUE
)
```

Or remove a file by exact path:

```{r, eval = FALSE}
pq_remove(file_name = company_archive)
```

## Related Pages

- Data management article
- WRDS to Parquet article
- PostgreSQL to Parquet article
- Parquet file utility reference pages
