---
title: "PostgreSQL to Parquet"
---

This package provides some basic functionality for exporting data from a PostgreSQL database to Parquet files.
Because the existence of "last updated" metadata cannot be assured with general PostgreSQL data sources, I have *not* implemented the `pg_update_pq()` function analogous to `wrds_update_pq()`.
Use the Python package for `pg_update_pq()`.

```{r}
library(db2pq)
```

## Database connections

The `db_to_pq()` function accepts arguments for `host`, `database`, `user`, `password`, and `port`.
The default values for these are taken from environment variables `PGHOST`, `PGDATABASE`, `PGUSER`, `PGPASSWORD`, and `PGPORT`, respectively.
Values for these arguments can also be given to `db_schema_to_pq()`, which will pass them along to the `db_to_pq()` calls it makes.

## Using `db_to_pq()`

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`db_to_pq()` exports PostgreSQL tables to Parquet using a DBI connection. It can
filter rows, select columns, rename outputs, and apply Arrow type overrides.

```{r, eval = FALSE}
db_to_pq(
  table_name = "company",
  schema = "comp",
  data_dir = "~/pq_data",
  keep = c("gvkey", "conm"),
  rename = c(conm = "company_name"),
  col_types = list(company_name = "string")
)
```

## Using other functions

Use `db_schema_tables()` to inspect available relations and `db_schema_to_pq()` to export an entire schema.

```{r}
db_schema_tables("ff")
```

```{r}
db_schema_to_pq("ff", data_dir = "~/pq_data")
```

## PostgreSQL numeric columns

With the default `numeric_mode = "decimal"`, table exports preserve bounded PostgreSQL `NUMERIC(p, s)` columns as Parquet decimal columns.

```{r}
db_to_pq(
  table_name = "dsi",
  schema = "crsp",
  data_dir = "~/pq_data"
)
```

```{r}
library(arrow)

read_parquet("~/pq_data/crsp/dsi.parquet", as_data_frame = FALSE)
```

Any PostgreSQL `NUMERIC` columns without precision and scale remain text because a
Parquet decimal type requires those bounds.

Use `numeric_mode = "text"` to retain exact text values for all detected
`NUMERIC` columns, `numeric_mode = "float64"` when double-precision output is
the intended representation, or `numeric_mode = "raw"` to keep the backend
default.

```{r}
db_to_pq(
  table_name = "dsi",
  schema = "crsp",
  data_dir = "~/pq_data",
  numeric_mode = "float64"
)
```

```{r}
read_parquet("~/pq_data/crsp/dsi.parquet", as_data_frame = FALSE)
```

While the R package does not implement general "update logic" of the WRDS-facing functions, table comments are retained as Parquet metadata.

```{r}
pq_last_modified("dsi", "crsp", data_dir = "~/pq_data")
```
