---
title: "PostgreSQL to Parquet"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`db_to_pq()` exports PostgreSQL tables to Parquet using a DBI connection. It can
filter rows, select columns, rename outputs, and apply Arrow type overrides.

```{r, eval = FALSE}
db_to_pq(
  table_name = "company",
  schema = "comp",
  data_dir = "~/pq_data",
  keep = c("gvkey", "conm"),
  rename = c(conm = "company_name"),
  col_types = list(company_name = "string")
)
```

Use `db_schema_tables()` to inspect available relations and `db_schema_to_pq()`
to export a schema.

## PostgreSQL Numeric Columns

By default, table exports preserve bounded PostgreSQL `NUMERIC(p, s)` columns
as Parquet decimal columns:

```{r, eval = FALSE}
db_to_pq(
  table_name = "dsi",
  schema = "crsp",
  data_dir = "~/pq_data"
)
```

The default `numeric_mode = "decimal"` detects `NUMERIC` columns from
PostgreSQL metadata, transfers them as text, and casts columns with known
precision and scale back to Arrow decimal types before writing Parquet.
PostgreSQL `NUMERIC` columns without precision and scale remain text because a
Parquet decimal type requires those bounds.

Use `numeric_mode = "text"` to retain exact text values for all detected
`NUMERIC` columns, `numeric_mode = "float64"` when double-precision output is
the intended representation, or `numeric_mode = "raw"` to keep the backend
default.
