---
title: "WRDS SSH Setup for SAS Metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WRDS SSH Setup for SAS Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Most `db2pq` WRDS work uses WRDS PostgreSQL credentials. SSH access is a
separate setup used only when `wrds_update_pq()` needs metadata from the WRDS
SAS side:

```{r, eval = FALSE}
wrds_update_pq("dsi", "crsp", use_sas = TRUE)
```

With `use_sas = TRUE`, `db2pq` retrieves last-modified dates by running
`PROC CONTENTS` on the WRDS SAS server over SSH. This can be useful for tables
whose PostgreSQL comment does not carry a reliable date.

## Before You Start

The `processx` package is required for the SAS metadata option:

```{r, eval = FALSE}
install.packages("processx")
```

Configure PostgreSQL credentials separately for ordinary WRDS table downloads.
See the [authentication article](authentication.html) for the `WRDS_ID`,
`.pgpass`, and `wrds::wrds_set_credentials()` paths supported by `db2pq`.

## Configure WRDS SSH

WRDS provides a dedicated SSH endpoint for key-based authentication:
`wrds-cloud-sshkey.wharton.upenn.edu`.

### Generate an SSH Key

The ed25519 key type is a reasonable default:

```sh
ssh-keygen -t ed25519 -C "your_wrds_id@wrds"
```

Accept the default location (`~/.ssh/id_ed25519`). You may use a passphrase if
your SSH agent is running. For unattended scripts, an empty passphrase may be
required.

### Copy the Public Key to WRDS

```sh
cat ~/.ssh/id_ed25519.pub | \
  ssh your_wrds_id@wrds-cloud-sshkey.wharton.upenn.edu \
  "mkdir -p ~/.ssh && chmod 700 ~/.ssh && \
   cat >> ~/.ssh/authorized_keys && chmod 600 ~/.ssh/authorized_keys"
```

### Configure SSH

Add an entry to `~/.ssh/config`:

```text
Host wrds
    HostName wrds-cloud-sshkey.wharton.upenn.edu
    User your_wrds_id
    IdentityFile ~/.ssh/id_ed25519
    IdentitiesOnly yes
```

Verify the connection:

```sh
ssh wrds
```

## Troubleshooting

If SSH still prompts for a password, inspect the connection:

```sh
ssh -vvv wrds
```

Confirm that `publickey` appears in the list of authentication methods. Also
check that `~/.ssh/config` uses the WRDS SSH endpoint and the key you copied to
WRDS.

## Related Pages

- [Authentication](authentication.html)
- [WRDS to Parquet](wrds-to-parquet.html)
