---
title: "WRDS to Parquet"
---

```{r}
library(db2pq)
```

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)
live_wrds_id <- tryCatch(db2pq::wrds_get_username(), error = function(e) "")
has_wrds_id <- nzchar(live_wrds_id)
has_libpq_password <- function(user) {
  nzchar(Sys.getenv("PGPASSWORD", unset = "")) ||
    tryCatch(
      db2pq:::pgpass_has_entry(
        host = "wrds-pgdata.wharton.upenn.edu",
        port = 9737L,
        database = "wrds",
        user = user
      ),
      error = function(e) FALSE
    )
}
live_wrds <- all(
  # identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_wrds_id,
  has_libpq_password(live_wrds_id),
  has_pkg("arrow"),
  has_pkg("dplyr"),
  has_pkg("DBI"),
  has_pkg("RPostgres")
)
old_data_dir <- Sys.getenv("DATA_DIR", unset = NA_character_)
doc_data_dir <- "db2pq-wrds-to-parquet-examples"
if (live_wrds) {
  unlink(doc_data_dir, recursive = TRUE)
  dir.create(doc_data_dir, recursive = TRUE, showWarnings = FALSE)
  Sys.setenv(DATA_DIR = doc_data_dir)
}
```

## Overview

The main helper on this path is `wrds_update_pq()`. 
It downloads a WRDS PostgreSQL table as a Parquet file and skips the download when the existing local file already appears current.

Related helpers are:

- `wrds_sql_to_pq()` exports a custom SQL query against WRDS into the standard
  Parquet layout.
- `wrds_schema_to_pq()` updates all or selected tables in a WRDS schema.
- `pq_last_modified()`, `pq_archive()`, and `pq_restore()` inspect and manage
  local Parquet files after they have been created.

## Using `wrds_update_pq()`

### An illustrative first run

A good table to start with is `comp.company`, which is small enough for testing:

```{r, eval = live_wrds}
wrds_update_pq("company", "comp")
```

If you run the function above and have not set up your details for a WRDS connection, you will be asked to provide these (e.g., your WRDS username and password).
More discussion on this topic can be found in the article on [authentication](authentication.html).

In addition, if you have not established `DATA_DIR`, as discussed in the article on [data management](data-management.html), you will be asked to provide that.


Running the same command later should usually skip the download if WRDS
metadata indicate that the local file is already current.

```{r live-company-second-run, eval = live_wrds}
company_file <- wrds_update_pq("company", "comp")
```

The resulting file is an ordinary Parquet file.
It can be inspected with `arrow`, `dplyr`, DuckDB, Python, or any other tool that reads Parquet.

```{r}
#| message: false
library(arrow)
library(dplyr, warn.conflicts = FALSE)
```

```{r live-company-preview, eval = live_wrds}
read_parquet(company_file) |> select(gvkey, conm, sic)
```

### Identifying Table Names and Schemas

WRDS uses schema names such as `crsp`, `comp`, `audit`, and `bank`. 

```{r, eval = FALSE}
wrds_update_pq("funda", "comp", obs = 1000)
```

You can list tables in a WRDS schema from R:

```{r, eval = live_wrds}
crsp_tables <- wrds_get_tables("crsp", views = TRUE)
crsp_tables[1:10]
```

Use the PostgreSQL schema name exposed by WRDS, which can differ from a short product name or SAS library name. 
For example, the Fama-French PostgreSQL base-table schema is `ff_all`, so a schema download starts with:

```{r, eval = live_wrds}
wrds_get_tables("ff_all")
```

However, these tables are exposed as **views** in the database schema `ff`.
Setting `views = TRUE` shows the underlying data made available there.

```{r, eval = live_wrds}
wrds_schema_to_pq("ff", views = TRUE)
```

The WRDS web query interface can also be useful for identifying the underlying
schema and table behind a dataset, but once the table is known, you should download with a `db2pq` script for reproducibility.

## Cleaning and Shaping Data

### Correcting Data Types

Some WRDS columns are stored in PostgreSQL with a type that is not ideal for
analysis. For example, link-history identifiers in `crsp.ccmxpf_lnkhist` are
often better represented as integers:

```{r, eval = live_wrds}
wrds_update_pq(
  "ccmxpf_lnkhist", "crsp",
  col_types = list(lpermno = "int32", lpermco = "int32")
)
```

### Setting Time Zones

WRDS PostgreSQL often stores timestamps without time-zone information. 
The `tz` argument tells `db2pq` how to interpret those values before writing normalized timestamps to Parquet.

For example, in the FFIEC Call Report data in the WRDS `bank` schema, timestamps should be
interpreted as US Eastern time:

```{r, eval = FALSE}
wrds_update_pq(
  "wrds_call_rcfa_1",
  "bank",
  tz = "America/New_York"
)
```

Some timestamp-like fields are stored as strings.
Use `col_types` with `tz` to coerce selected fields:

```{r, eval = FALSE}
wrds_update_pq(
  "feed03_audit_fees",
  "audit",
  keep = c("auditor_fkey", "file_accepted"),
  col_types = list(
    auditor_fkey = "int32",
    file_accepted = "timestamp"
  ),
  tz = "America/New_York",
  force = TRUE
)
```

## Working With Large Tables

### Dropping Unwanted Variables

Some WRDS tables contain blocks of columns that are not needed for most analyses. 
The `drop` argument removes columns whose names match one or more regular expressions:

```{r, eval = FALSE}
wrds_update_pq(
  "feed02_auditor_changes",
  "audit",
  drop = "^(match|prior|closest)"
)
```

### Selecting Only Certain Variables

Use `keep` when you want a small subset of columns:

```{r, eval = live_wrds}
dsf_file = wrds_update_pq(
  "dsf",
  "crsp",
  obs = 100,
  force = TRUE,
  keep = c("permno", "date", "^ret$")
)

read_parquet(dsf_file)
```

Note that the arguments to `keep` are evaluated as **regular expressions**, so you want to use `"^ret$"` to match `ret` rather than any column *containing* `ret`.
For more on regular expressions, see [Chapter 9](https://iangow.github.io/far_book/web-data.html) of *Empirical Research in Accounting: Tools and Methods*.

### Renaming Variables

Use `rename` to write output columns with preferred names:

```{r, eval = FALSE}
wrds_update_pq(
  "company",
  "comp",
  keep = c("gvkey", "conm", "sic"),
  alt_table_name = "sic_codes",
  rename = c(conm = "company_name")
)
```

For a small documentation example, the renamed file can be written to the temporary repository:

```{r live-company-rename, eval = live_wrds}
renamed_company_file <- wrds_update_pq(
  "company",
  "comp",
  alt_table_name = "company_names",
  keep = c("gvkey", "conm", "sic"),
  rename = c(conm = "company_name"),
  force = TRUE
)

read_parquet(renamed_company_file) |>
  select(gvkey, company_name, sic)
```

If `rename` and `col_types` are used together, names in `col_types` refer to
the output names after renaming:

```{r, eval = FALSE}
wrds_update_pq(
  "ccmxpf_lnkhist",
  "crsp",
  rename = c(lpermno = "permno_link"),
  col_types = list(permno_link = "int32")
)
```

### Selecting Only Certain Rows

Use `where` for SQL row filters. For Compustat annual fundamentals, a common
filter is the standard industrial format, standard data format, consolidated,
domestic-population set:

```{r, eval = FALSE}
wrds_update_pq(
  "funda",
  "comp",
  where = paste(
    "indfmt = 'INDL'",
    "AND datafmt = 'STD'",
    "AND consol = 'C'",
    "AND popsrc = 'D'"
  )
)
```

### Limiting Rows for a Test Download

Use `obs` to test a download before running the full export:

```{r, eval = FALSE}
wrds_update_pq("funda", "comp", obs = 1000)
```

Combining `keep`, `where`, and `obs` is a useful way to produce a small local
file while developing a refresh script.

```{r live-funda-sample, eval = live_wrds}
funda_sample_file <- wrds_update_pq(
  "funda",
  "comp",
  alt_table_name = "funda_sample",
  keep = c("gvkey", "datadate", "fyear", "^at$", "^sale$"),
  where = paste(
    "indfmt = 'INDL'",
    "AND datafmt = 'STD'",
    "AND consol = 'C'",
    "AND popsrc = 'D'"
  ),
  obs = 1000,
  force = TRUE
)

read_parquet(funda_sample_file) |> glimpse()
```

## Update Logic

### Forcing an Update

Use `force = TRUE` when you want to re-download even if local metadata suggest
the file is current. This is useful after changing `keep`, `drop`, `rename`,
`col_types`, or `tz`.

```{r, eval = FALSE}
wrds_update_pq("funda", "comp", force = TRUE, obs = 1000)
```

### Using SAS Metadata

By default, `wrds_update_pq()` compares the local Parquet metadata with WRDS
PostgreSQL table comments. Some WRDS PostgreSQL tables do not have useful
comments. In those cases, `use_sas = TRUE` can retrieve update metadata from
the corresponding WRDS SAS dataset.

Some schemas have different names on the PostgreSQL and SAS sides. For example,
RavenPack's PostgreSQL schema can differ from its SAS library:

```{r, eval = FALSE}
wrds_update_pq(
  "rpa_entity_mappings",
  "ravenpack_common",
  use_sas = TRUE,
  sas_schema = "rpa"
)
```

This path requires SSH access to the WRDS SAS server. See the
[WRDS SSH setup article](wrds-ssh.html).

### Archiving Data When Updating

Use `archive = TRUE` when you want replaced files moved into the schema's
archive directory before the new file is written:

```{r, eval = FALSE}
wrds_update_pq("company", "comp", archive = TRUE)
pq_last_modified("company", "comp", archive = TRUE)
```

Archives live under:

```text
<DATA_DIR>/<schema>/archive/
```

```{r live-company-archive, eval = live_wrds}
archived_company <- pq_archive("company", "comp")
basename(archived_company)

pq_last_modified("company", "comp", archive = TRUE) |>
  select(file_name, last_mod)
```

The archived copy can be restored, again without touching the main `DATA_DIR`.

```{r live-company-restore, eval = live_wrds}
archived_company_name <- basename(archived_company)

pq_restore(
  archived_company_name,
  "comp",
  archive = FALSE
)

pq_last_modified("company", "comp")
```

## Custom SQL

Use `wrds_sql_to_pq()` when the object you want is better expressed as a query
than as a full table export:

```{r, eval = FALSE}
wrds_sql_to_pq(
  "SELECT permno, date, ret FROM crsp.dsf WHERE date >= '2024-01-01'",
  table_name = "dsf_recent",
  schema = "crsp"
)
```

For a small query result, use `LIMIT`:

```{r live-sql-export, eval = live_wrds}
sql_query <- "
    SELECT gvkey, datadate, fyear, at, sale
    FROM comp.funda
    WHERE indfmt = 'INDL'
    AND datafmt = 'STD'
    AND consol = 'C'
    AND popsrc = 'D'
    ORDER BY gvkey, datadate
    LIMIT 20
  "

sql_file <- wrds_sql_to_pq(
  sql_query,
  table_name = "funda_sql_sample",
  schema = "comp",
  archive = TRUE
)

read_parquet(sql_file)
```

## Related Pages

- `wrds_update_pq()`
- `wrds_sql_to_pq()`
- `wrds_schema_to_pq()`
- `pq_last_modified()`
- `pq_archive()`
- Authentication article
- Data management article

```{r, include = FALSE}
if (live_wrds) {
  unlink(doc_data_dir, recursive = TRUE)
  if (is.na(old_data_dir)) {
    Sys.unsetenv("DATA_DIR")
  } else {
    Sys.setenv(DATA_DIR = old_data_dir)
  }
}
```
