---
title: "WRDS to Parquet"
---

```{r}
library(db2pq)
```

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_pkg <- function(pkg) requireNamespace(pkg, quietly = TRUE)
live_wrds_id <- tryCatch(db2pq::wrds_get_username(), error = function(e) "")
has_wrds_id <- nzchar(live_wrds_id)
has_libpq_password <- function(user) {
  nzchar(Sys.getenv("PGPASSWORD", unset = "")) ||
    tryCatch(
      db2pq::pgpass_has_entry(
        host = "wrds-pgdata.wharton.upenn.edu",
        port = 9737L,
        database = "wrds",
        user = user
      ),
      error = function(e) FALSE
    )
}
live_wrds <- all(
  identical(Sys.getenv("IN_PKGDOWN"), "true"),
  has_wrds_id,
  has_libpq_password(live_wrds_id),
  has_pkg("arrow"),
  has_pkg("dplyr"),
  has_pkg("DBI"),
  has_pkg("RPostgres")
)
doc_data_dir <- file.path(tempdir(), "db2pq-wrds-to-parquet")
if (live_wrds) {
  unlink(doc_data_dir, recursive = TRUE)
  dir.create(doc_data_dir, recursive = TRUE, showWarnings = FALSE)
}
```

This is the place to start when your source data live in the WRDS PostgreSQL
database and you want local Parquet files that can be reused across projects.

`db2pq` writes files using the layout:

```text
<DATA_DIR>/<schema>/<table>.parquet
```

For example, `wrds_update_pq("company", "comp")` writes
`<DATA_DIR>/comp/company.parquet` unless you provide `data_dir` or `out_file`
explicitly.

## Main Helpers

The main helper on this path is `wrds_update_pq()`. It downloads a WRDS
PostgreSQL table as a Parquet file and skips the download when the existing
local file already appears current.

Related helpers are:

- `wrds_sql_to_pq()` exports a custom SQL query against WRDS into the standard
  Parquet layout.
- `wrds_schema_to_pq()` updates all or selected tables in a WRDS schema.
- `pq_last_modified()`, `pq_archive()`, and `pq_restore()` inspect and manage
  local Parquet files after they have been created.

## Setup

You need to set up your details for a WRDS connection as discussed in the article on [authentication article](authentication.html).

Set `DATA_DIR` to the local repository where you want Parquet files to live:

```{r, eval = FALSE}
Sys.setenv(DATA_DIR = "~/Dropbox/pq_data")
```


## Using `wrds_update_pq()`

### An illustrative first run

A good first table is `comp.company`, which is small enough for a quick test:

```{r, eval = FALSE}
wrds_update_pq("company", "comp")
```

When this article is rendered locally with WRDS credentials, the same example
runs against a temporary documentation repository rather than the main
`DATA_DIR`.

```{r live-company-first-run, eval = live_wrds}
library(db2pq)

wrds_id <- wrds_get_username()
company_file <- wrds_update_pq(
  "company",
  "comp",
  data_dir = doc_data_dir,
  wrds_id = wrds_id
)
file.path("<temporary data repository>", "comp", basename(company_file))
```

Typical output from a first run looks like:

```text
Updated comp.company is available.
Beginning file download at 2026-04-09 17:12:41 UTC.
Completed file download at 2026-04-09 17:12:46 UTC.
```

Running the same command later should usually skip the download if WRDS
metadata indicate that the local file is already current.

```{r live-company-second-run, eval = live_wrds}
wrds_update_pq("company", "comp", data_dir = doc_data_dir, wrds_id = wrds_id)
```

The resulting file is an ordinary Parquet file. It can be inspected with
`arrow`, `dplyr`, DuckDB, Python, or any other tool that reads Parquet.

```{r live-company-preview, eval = live_wrds}
dplyr::select(
  arrow::read_parquet(company_file),
  gvkey, conm, sic
) |>
  head()
```

### Identifying Table Names and Schemas

WRDS uses schema names such as `crsp`, `comp`, `audit`, and `bank`. If existing
code refers to `comp.funda`, then the schema is `comp` and the table is
`funda`:

```{r, eval = FALSE}
wrds_update_pq("funda", "comp")
```

You can list tables in a WRDS schema from R:

```{r, eval = FALSE}
wrds_get_tables("crsp")
wrds_get_tables("comp", views = TRUE)
```

Use the PostgreSQL schema name exposed by WRDS, which can differ from a short
product name or SAS library name. For example, the Fama-French PostgreSQL
base-table schema is `ff_all`, so a schema download starts with:

```{r, eval = FALSE}
wrds_schema_to_pq("ff_all")
```

Schemas containing views need `views = TRUE` during discovery. For example,
relations exposed through the Fama-French `ff` schema can be processed with:

```{r, eval = FALSE}
wrds_schema_to_pq("ff", views = TRUE)
```

During a local site render, this page can show the first few tables returned
from WRDS:

```{r live-list-comp-tables, eval = live_wrds}
head(wrds_get_tables("comp", wrds_id = wrds_id))
```

The WRDS web query interface can also be useful for identifying the underlying
schema and table behind a dataset, but once the table is known, script the
download with `db2pq` so it is reproducible.

## Cleaning and Shaping Data

### Correcting Data Types

Some WRDS columns are stored in PostgreSQL with a type that is not ideal for
analysis. For example, link-history identifiers in `crsp.ccmxpf_lnkhist` are
often better represented as integers:

```{r, eval = FALSE}
wrds_update_pq(
  "ccmxpf_lnkhist",
  "crsp",
  col_types = list(lpermno = "int32", lpermco = "int32")
)
```

### Setting Time Zones

WRDS PostgreSQL often stores timestamps without time-zone information. The `tz`
argument tells `db2pq` how to interpret those values before writing normalized
timestamps to Parquet.

For FFIEC Call Report data in the WRDS `bank` schema, timestamps should be
interpreted as US Eastern time:

```{r, eval = FALSE}
wrds_update_pq(
  "wrds_call_rcfa_1",
  "bank",
  tz = "America/New_York"
)
```

Some timestamp-like fields are stored as strings. Use `col_types` with `tz` to
coerce selected fields:

```{r, eval = FALSE}
wrds_update_pq(
  "feed03_audit_fees",
  "audit",
  keep = c("auditor_fkey", "file_accepted"),
  col_types = list(
    auditor_fkey = "int32",
    file_accepted = "timestamp"
  ),
  tz = "America/New_York",
  force = TRUE
)
```

## Working With Large Tables

### Dropping Unwanted Variables

Some WRDS tables contain blocks of columns that are not needed for most
analyses. The `drop` argument removes columns whose names match one or more
regular expressions:

```{r, eval = FALSE}
wrds_update_pq(
  "feed02_auditor_changes",
  "audit",
  drop = "^(match|prior|closest)"
)
```

### Selecting Only Certain Variables

Use `keep` when you want a small subset of columns:

```{r, eval = FALSE}
wrds_update_pq(
  "dsf",
  "crsp",
  keep = c("permno", "date", "ret")
)
```

### Renaming Variables

Use `rename` to write output columns with preferred names:

```{r, eval = FALSE}
wrds_update_pq(
  "company",
  "comp",
  keep = c("gvkey", "conm", "sic"),
  rename = c(conm = "company_name")
)
```

For a small documentation example, the renamed file can be written to the
temporary repository:

```{r live-company-rename, eval = live_wrds}
renamed_company_file <- wrds_update_pq(
  "company",
  "comp",
  data_dir = doc_data_dir,
  alt_table_name = "company_names",
  keep = c("gvkey", "conm", "sic"),
  rename = c(conm = "company_name"),
  wrds_id = wrds_id,
  force = TRUE
)

dplyr::select(
  arrow::read_parquet(renamed_company_file),
  gvkey, company_name, sic
) |>
  head()
```

If `rename` and `col_types` are used together, names in `col_types` refer to
the output names after renaming:

```{r, eval = FALSE}
wrds_update_pq(
  "ccmxpf_lnkhist",
  "crsp",
  rename = c(lpermno = "permno_link"),
  col_types = list(permno_link = "int32")
)
```

### Selecting Only Certain Rows

Use `where` for SQL row filters. For Compustat annual fundamentals, a common
filter is the standard industrial format, standard data format, consolidated,
domestic-population set:

```{r, eval = FALSE}
wrds_update_pq(
  "funda",
  "comp",
  where = paste(
    "indfmt = 'INDL'",
    "AND datafmt = 'STD'",
    "AND consol = 'C'",
    "AND popsrc = 'D'"
  )
)
```

### Limiting Rows for a Test Download

Use `obs` to test a download before running the full export:

```{r, eval = FALSE}
wrds_update_pq("funda", "comp", obs = 1000)
```

Combining `keep`, `where`, and `obs` is a useful way to produce a small local
file while developing a refresh script:

```{r live-funda-sample, eval = live_wrds}
funda_sample_file <- wrds_update_pq(
  "funda",
  "comp",
  data_dir = doc_data_dir,
  alt_table_name = "funda_sample",
  keep = c("gvkey", "datadate", "fyear", "at", "sale"),
  where = paste(
    "indfmt = 'INDL'",
    "AND datafmt = 'STD'",
    "AND consol = 'C'",
    "AND popsrc = 'D'"
  ),
  obs = 1000,
  wrds_id = wrds_id,
  force = TRUE
)

dplyr::glimpse(arrow::read_parquet(funda_sample_file))
```

## Update Logic

### Forcing an Update

Use `force = TRUE` when you want to re-download even if local metadata suggest
the file is current. This is useful after changing `keep`, `drop`, `rename`,
`col_types`, or `tz`.

```{r, eval = FALSE}
wrds_update_pq("funda", "comp", force = TRUE)
```

### Using SAS Metadata

By default, `wrds_update_pq()` compares the local Parquet metadata with WRDS
PostgreSQL table comments. Some WRDS PostgreSQL tables do not have useful
comments. In those cases, `use_sas = TRUE` can retrieve update metadata from
the corresponding WRDS SAS dataset.

Some schemas have different names on the PostgreSQL and SAS sides. For example,
RavenPack's PostgreSQL schema can differ from its SAS library:

```{r, eval = FALSE}
wrds_update_pq(
  "rpa_entity_mappings",
  "ravenpack_common",
  use_sas = TRUE,
  sas_schema = "rpa"
)
```

This path requires SSH access to the WRDS SAS server. See the
[WRDS SSH setup article](wrds-ssh.html).

### Archiving Data When Updating

Use `archive = TRUE` when you want replaced files moved into the schema's
archive directory before the new file is written:

```{r, eval = FALSE}
wrds_update_pq("company", "comp", archive = TRUE)
pq_last_modified("company", "comp", archive = TRUE)
```

Archives live under:

```text
<DATA_DIR>/<schema>/archive/
```

For documentation and testing, archive examples should use a temporary data
repository. This keeps the real local Parquet repository unchanged while still
exercising the same file-management code.

```{r live-company-archive, eval = live_wrds, message = FALSE}
archived_company <- pq_archive("company", "comp", data_dir = doc_data_dir)
basename(archived_company)

pq_last_modified("company", "comp", data_dir = doc_data_dir, archive = TRUE) |>
  dplyr::select(file_name, last_mod)
```

The archived copy can be restored, again without touching the main `DATA_DIR`.

```{r live-company-restore, eval = live_wrds, message = FALSE}
pq_restore(
  tools::file_path_sans_ext(basename(archived_company)),
  "comp",
  data_dir = doc_data_dir,
  archive = FALSE
)

pq_last_modified("company", "comp", data_dir = doc_data_dir)
```

## Custom SQL

Use `wrds_sql_to_pq()` when the object you want is better expressed as a query
than as a full table export:

```{r, eval = FALSE}
wrds_sql_to_pq(
  "SELECT permno, date, ret FROM crsp.dsf WHERE date >= '2024-01-01'",
  table_name = "dsf_recent",
  schema = "crsp"
)
```

For a small query result, use `LIMIT` and write to the temporary documentation
repository:

```{r live-sql-export, eval = live_wrds}
sql_file <- wrds_sql_to_pq(
  paste(
    "SELECT gvkey, datadate, fyear, at, sale",
    "FROM comp.funda",
    "WHERE indfmt = 'INDL'",
    "AND datafmt = 'STD'",
    "AND consol = 'C'",
    "AND popsrc = 'D'",
    "ORDER BY gvkey, datadate",
    "LIMIT 20"
  ),
  table_name = "funda_sql_sample",
  schema = "comp",
  data_dir = doc_data_dir,
  wrds_id = wrds_id,
  archive = TRUE
)

arrow::read_parquet(sql_file)
```

## Related Pages

- `wrds_update_pq()`
- `wrds_sql_to_pq()`
- `wrds_schema_to_pq()`
- `pq_last_modified()`
- `pq_archive()`
- Authentication article
- Data management article
